---
title: "Groundhog"
description: |
  Analysis of Groundhog Prediction data from TidyTuesday
author: Devi Canseco
date: February 18, 2025
format: html
---

```{r, include=FALSE}

tuesdata <- tidytuesdayR::tt_load('2024-01-30')

groundhogs <- tuesdata$groundhogs
predictions <- tuesdata$predictions

library(tidyverse)
```

**Introduction**

This project draws its data from the GitHub repository TidyTuesday's 2024-01-30 folder titled "Groundhog Predictions." This data is originally from https://groundhog-day.com/. This data puts together the predictions of groundhogs for 6 more weeks of winter or the coming of spring across the US and Canada. That is to say, several groundhogs' predictions are recorded every year.

```{r}
shadow_prediction  <- predictions |>
  group_by(year) |>
  summarize(total_shadow_count = sum(shadow, na.rm = TRUE))

ggplot(shadow_prediction, aes(x = year, y = total_shadow_count)) +
  geom_point()


```

**Graphing and Code**

```{r}
id_prediction  <- predictions |>
  group_by(id) |>
  summarize(total_shadow_count = sum(shadow, na.rm = TRUE))

ggplot(id_prediction,  aes(x = id, y = total_shadow_count)) +
  geom_col() +
  labs(
    title = "Longer winter predictions per Ground Hog",
    x = "Groundhog identification",
    y = "Saw Shadow Count"
  )

ggplot(predictions, aes(x = year, fill = shadow)) +
  geom_bar() +
  labs(
    title = "Yearly Shadow Predictions",
    x = "Year",
    y = "Total Predictions"
  )
```

**Analysis**

The first graph isolates the total count of observed shadows by year. Similarly, the third graph reiterates this idea by presenting the count of all shadow predictions per year. From these graphs, it is obvious that there is a growing number of predictions by different groundhogs. The growth in groundhog day popularity is especially visible when we view the third graph in conjunction with the second graph showing the consistency and recency in groundhog count of predictions. In the second graph, we can see that earlier groundhogs have had a greater impact on the data for winter predictions.
