---
title: "Simpsons's Script"
description: |
  Analysis of Simpsons data from TidyTuesday
author: Devi Canseco
date: February 18, 2025
format: html
---

**Introduction**

In this project, I will use data from the TidyTuesday folder for February 2, 2025. The original data is from <https://www.kaggle.com/datasets/prashant111/the-simpsons-dataset>, which is a kraggle data set authored by Prashant Banerjee. This data is further scraped by other users. The data is split into four data frames of characters, episodes, locations, and script lines. I focused on script line count among the characters with the over 2000 cumulative word counts.

```{r, include=FALSE}

tuesdata <- tidytuesdayR::tt_load('2025-02-04')

simpsons_characters <- tuesdata$simpsons_characters
simpsons_episodes <- tuesdata$simpsons_episodes
simpsons_locations <- tuesdata$simpsons_locations
simpsons_script_lines <- tuesdata$simpsons_script_lines

library(tidyverse)
```

**Graph and Code**

```{r}
known_character <- simpsons_script_lines |>
  group_by(raw_character_text) |>
  summarize(total_word_count = sum(word_count, na.rm = TRUE)) |>
  filter(total_word_count > 2000)|>
  arrange(total_word_count)

```

```{r}
  ggplot(known_character, aes(y = reorder(raw_character_text, -total_word_count), x = total_word_count)) +
  geom_col() +
  labs(
    title = "Characters with the Highest Word Count",
    y = "Characters by Name",
    x = "Total Word Count"
  ) + theme_minimal()

```

**Analysis**

For this data, we can see that the characters that have said over 2000 words are limited to 16 characters. Characters in the main cast naturally have more lines, but there is a large difference between the character with the highest word count (Homer) and the character with the second highest word count (Marge). This can be attributed to Homer being the protagonist. When we exclude the Simpsons family, Moe is the character with the most words said. There is a notable drop between Moe and Lisa Simpson, signaling the end of the main cast. Would an avid Simpsons fan consider this character's word count surprising?
