---
title: "English Education in Small Towns"
description: |
  Analysis of English Education data from TidyTuesday
author: Devi Canseco
date: April 4, 2025
format: html
---

```{r, inlcude = FALSE}
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load(2024, week = 4)

english_education <- tuesdata$english_education
```

In this analysis, I will utilize the data from TidyTuesday 2024-01-23 to perform a permutation test. The goal of the test is to confirm the claims of the article in which the data was used to answer the question: Is there a significant difference in education scores between different-sized towns? The data is taken from the UK Office for National Statistics.

Link to the data:

<https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-01-23/readme.md>

<https://www.ons.gov.uk/>

```{r}
size_interest <- english_education |>
  filter(size_flag == "Medium Towns"| size_flag =="Small Towns", na.rm = TRUE)
```

The selected variable

**Single permutation**

```{r}
permutation_test <- function(reps,data) {
  data |>
    select(size_flag, education_score)|>
    mutate(edu_perm = sample(education_score, replace = FALSE)) |>
    group_by(size_flag) |>
    summarize(mean_score = mean(education_score, na.rm = TRUE), mean_perm = mean(edu_perm)) |>
    summarize(mean_diff = abs(diff(mean_score)), mean_perm_diff = diff(mean_perm))
  }

```

**Repeated simulated trials**

```{r}
set.seed(123)
under_null <- map(1:10000, permutation_test, data = size_interest) |>
  list_rbind()
```

**Plot distribution**

```{r}
ggplot(under_null, aes(x = mean_perm_diff)) +
  geom_histogram(binwidth = 0.1, color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_diff, color = "red")) +
  labs(title = "Distribution of Permutation Under Null",
       x = "Difference in Mean Scores",
       y = "Frequency") 
```

From the plot it can be seen that the line representing the observational measurement of education score differences is off-centered. There is a small number of occurrences of simulated under the null. This would mean that the observed difference is more extreme than data created randomly.

```{r}
under_null|>
summarize(p_value = mean(mean_perm_diff > mean_diff))
```

After creating 10000 permutation simulations, the distribution under the null hypothesis falls on zero as there is no expected difference between the two groups. From the results of the p-value, if the significance level, as it usually is, is set to be 0.05, our p-value fall below this level. $\alpha$ \> 0.0139. This means that our data is significantly extreme and would probably not occur by chance.
