---
title: "SQL"
description: |
  Using SQL to analyze police traffic stops
  
author: Devi Canseco
date: April 15, 2025
format: html
---

```{r, include = FALSE}
library(DBI)
library(RMariaDB)
library(ggplot2)
library(tidyverse)

con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)

DBI::dbListTables(con_traffic)
```

The data for this analysis comes from the Stanford Open Policing Project, published in Pierson et al. (2020). Which can be found in <https://openpolicing.stanford.edu/data/>. In this project, I want to compare reasonable driving stops for drivers across states. First, I will compare drivers from California and Florida with the tentative question: Which state has better drivers? To answer this question, I will find and compare the proportion of stops that resulted in a citation for each state. On a second analysis, I will utilize Philadelphia's data frame to find that searches provided resulted in a reasonable cause when performed. This follows the question: Did they have the right suspicion? I will extract the number of searches that found contraband and compare it to the total number of searches conducted.

**Better Drivers**

In this analysis, the statewide reports of LA and FL are used to form a comparison of the proportion of drivers issued a citation. Since the coverage rate for state wide reports did not reach 100%, 69.8% and % for California and Florida, respectively, two cities in these states were chose as representatives of the state. Thereby we compare San Francisco drivers with Tampa drivers. Two cities with large populations.

```{sql}
#| connection: con_traffic
#| output.var: "prop_table"

SELECT
    state,
    COUNT(*) AS total_stops,
    SUM(citation_issued) AS citations,
    ROUND(SUM(citation_issued) / COUNT(*), 2) AS citation_rate
FROM (
    SELECT
        'CA' AS state,
         citation_issued
    FROM ca_san_francisco_2020_04_01

    UNION ALL

    SELECT
        'FL' AS state,
        citation_issued
    FROM fl_tampa_2020_04_01
) combined
GROUP BY state
ORDER BY citation_rate DESC;
```

```{r}
head(prop_table)
```

The table shows numerically how many stops and how many of those stops resulted in citations.

```{r}
ggplot(prop_table, aes(x = state, y = citation_rate)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportion of Drivers Issued Citations by State Patrol",
    x = "State",
    y = "Citation Rate"
  ) 
```

From the graph, it can be seen that there is a large difference between the proportion of drivers who were stopped AND received a citation in San Francisco, California and Tampa, Florida. While no conclusions can be drawn from this snippet of data, it can be speculated that perhaps Californian drivers are better drivers and did not commit violations that guaranteed a citation, or it could also propose the idea that traffic officers are more lenient to driving mistakes. Both show a small proportion, which could speak about the cities themselves.

**Evidence Found**

In this analysis, the data is taken from the Philadelphia, PA table. It seeks to find if searches resulted in finding evidence for a suspected holding of contraband.

```{sql}
#| connection: con_traffic
#| output.var: "contr_table"
SELECT
    COUNT(*) AS total_searches,
    ROUND(SUM(CASE WHEN contraband_found = TRUE THEN 1 ELSE 0 END)) AS searches_with_contraband,
    ROUND(SUM(CASE WHEN contraband_found = TRUE THEN 1 ELSE 0 END) / COUNT(*), 2) AS contraband_hit_rate 
FROM
    pa_philadelphia_2020_04_01
WHERE
    search_conducted = TRUE
    AND contraband_found IS NOT NULL;
```

```{r}
contr_table
```

```{r}
search_table <- contr_table |>
  mutate(searches_without_contraband = total_searches - searches_with_contraband)|>
  mutate(contraband_miss_rate = 100 - contraband_hit_rate)


search_table <- contr_table |>
  mutate(
    searches_without_contraband = total_searches - searches_with_contraband
  ) |>
  transmute(
    `Contraband Found` = as.numeric(searches_with_contraband),
    `No Contraband Found` = as.numeric(searches_without_contraband)
  ) |>
  pivot_longer(
    cols = everything(),
    names_to = "category",
    values_to = "count"
  )

ggplot(search_table, aes(x = category, y = count, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Philadelphia Search Outcomes",
    y = "Number of Searches",
    x = ""
  )

```

```{r}
DBI::dbDisconnect(con_traffic)
```

Data Source: Stanford Open Policing Project. Pierson, E., Corbett-Davies, S., & Goel, S. (2020). “A large-scale analysis of racial disparities in police stops across the United States.” *Nature Human Behaviour*, 4(7), 736–745. <https://openpolicing.stanford.edu/>
