---
title: "DS002R Projects Presentation"
author: "Devi Canseco"
subtitle: "May 13, 2025"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

# TidyTuesday

## Groundhog Day Predictions

-   TidyTuesday's January 30, 2024
-   Looking at the spring arrival predictions from groundhogs across America and Canada
-   Have predictions increased over the years, and do they lean towards one prediction result?
-   Which groundhog is contributing to the results?

## Increasing Predictions

```{r, echo =FALSE}
tuesdata <- tidytuesdayR::tt_load('2024-01-30')
groundhogs <- tuesdata$groundhogs
predictions <- tuesdata$predictions

library(tidyverse)

ggplot(predictions, aes(x = year, fill = shadow)) +
  geom_bar() +
  labs(
    title = "Yearly Shadow Predictions",
    x = "Year",
    y = "Total Predictions"
  ) + theme_minimal()

```

## Groundhog Contribution

```{r, echo=FALSE}
id_prediction  <- predictions |>
  group_by(id) |>
  summarize(total_shadow_count = sum(shadow, na.rm = TRUE))

ggplot(id_prediction,  aes(x = id, y = total_shadow_count)) +
  geom_col() +
  labs(
    title = "Longer winter predictions per Ground Hog",
    x = "Groundhog identification",
    y = "Saw Shadow Count"
  ) + theme_minimal()
```

## The Simpsons's Script

-   TidyTuesday folder for February 2, 2025

-   Looking at characters with over 2000 cumulative word counts.

-   How is line distribution decided based on a character's role?

## Character word count

```{r, echo=FALSE}
tuesdata <- tidytuesdayR::tt_load('2025-02-04')

simpsons_characters <- tuesdata$simpsons_characters
simpsons_episodes <- tuesdata$simpsons_episodes
simpsons_locations <- tuesdata$simpsons_locations
simpsons_script_lines <- tuesdata$simpsons_script_lines

library(tidyverse)

known_character <- simpsons_script_lines |>
  group_by(raw_character_text) |>
  summarize(total_word_count = sum(word_count, na.rm = TRUE)) |>
  filter(total_word_count > 2000)|>
  arrange(total_word_count)

ggplot(known_character, aes(y = reorder(raw_character_text, -total_word_count), x = total_word_count)) +
  geom_col() +
  labs(
    title = "Characters with the Highest Word Count",
    y = "Characters by Name",
    x = "Total Word Count"
  ) + theme_minimal()
```

## Conclusions

-   Main character, main cast, popular side characters

-   Would the results be surprising to a fan? Can it be used to gain insight into popularity?

# Risk Profile

-   TidyTuesday January 23, 2024

-   Comparing education scores in medium-sized vs small-sized towns

-   Is there a significant difference in education scores between different-sized towns?

## Repeated Permutation

```{r, echo=FALSE}
tuesdata <- tidytuesdayR::tt_load(2024, week = 4)
english_education <- tuesdata$english_education

size_interest <- english_education |>
  filter(size_flag == "Medium Towns"| size_flag =="Small Towns", na.rm = TRUE)

permutation_test <- function(reps,data) {
  data |>
    select(size_flag, education_score)|>
    mutate(edu_perm = sample(education_score, replace = FALSE)) |>
    group_by(size_flag) |>
    summarize(mean_score = mean(education_score, na.rm = TRUE), mean_perm = mean(edu_perm)) |>
    summarize(mean_diff = abs(diff(mean_score)), mean_perm_diff = diff(mean_perm))
}

set.seed(123)
under_null <- map(1:10000, permutation_test, data = size_interest) |>
  list_rbind()

ggplot(under_null, aes(x = mean_perm_diff)) +
  geom_histogram(binwidth = 0.1, color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean_diff, color = "red")) +
  labs(title = "Distribution of Permutation Under Null",
       x = "Difference in Mean Scores",
       y = "Frequency") +
  theme_minimal()
```

## Conclusions

-   P-value = 0.0139

-   There is a significant difference between education scores of medium-sized and small-sized towns

# SQL Drivers

## Florida vs California

-   Stanford Open Policing Project

-   Comparing Florida's and California's proportion of total citations issued per total stops.

-   Which state has better drivers?

## Proportion

```{r, echo=FALSE}

 library(DBI)
library(RMariaDB)
library(ggplot2)
library(tidyverse)

con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)

```

```{sql, echo=FALSE}
#| connection: con_traffic
#| output.var: "prop_table"

SELECT
    state,
    COUNT(*) AS total_stops,
    SUM(citation_issued) AS citations,
    ROUND(SUM(citation_issued) / COUNT(*), 2) AS citation_rate
FROM (
    SELECT
        'CA' AS state,
         citation_issued
    FROM ca_statewide_2023_01_26

    UNION ALL

    SELECT
        'FL' AS state,
        citation_issued
    FROM fl_statewide_2020_04_01
) combined
GROUP BY state
ORDER BY citation_rate DESC;
```

```{r, echo=FALSE}
ggplot(prop_table, aes(x = state, y = citation_rate)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportion of Drivers Issued Citations by State Patrol",
    x = "State",
    y = "Citation Rate"
  ) 
DBI::dbDisconnect(con_traffic)
```

## Conclusions

-   Not a determinant of good drivers, but fun speculation of the causes for the difference.

-   Skill of the drivers or police attitude?
